### 2023-12-27 sections/2-methods.tex
misc minor changes

We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and {\color{red} have an identity (TODO: improve the wording for this phrase)} (i.e. $f(x, y) = 0 \iff x = y$).
If, in addition to these constraints, the distance function obeys the triangle inequality (i.e. $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then the distance function is also a \emph{distance metric}.
Under distance metrics, all search algorithms in CAKES are exact.
For example, Euclidean, Levenshtein and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality.

@@

% Najib: For this figure, I want the red, green and blue lines/braces to be thinner. Perhaps 1/2 to 2/3 as thick. Also, I changed the notation for the deltas because it makes things easier in the Sieve and Greedy Sieve algorithms' sections.
\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
    \caption{$\delta$, $\delta^{+}$, and $\delta^{-}$ for a cluster $C$ and a query $q$.}
    \label{fig:methods:deltas}
\end{figure}

@@

% Najib: Visually, ``m'' is not centered on the line segment. It's a little bit to the right. Moving ``l'' to the right is probably the easiest fix.
\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.5]{images/geometry/overlapping-children-3.pdf}
    \caption{
        $m$ denotes the midpoint of $\overline{l r}$. $q'$ is the projection of $q$ onto $\overline{l r}$.
        $d$ is the distance from $q'$ to $m$.
        If $d \leq \rho$, both child clusters could contain points in the query ball.
        Otherwise, only the right child cluster could contain such points. 
    }
    \label{fig:methods:overlapping-children-3}
\end{figure}

>>>>

We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and has identity of indiscernibles (i.e. $f(x, y) = 0 \iff x = y$).
If, in addition to these constraints, the distance function obeys the triangle inequality (i.e. $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then the distance function is also a \emph{distance metric}.
Under distance metrics, all search algorithms in CAKES are exact.
For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
    \caption{$\delta$, $\delta^{+}$, and $\delta^{-}$ for a cluster $C$ and a query $q$.}
    \label{fig:methods:deltas}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.5]{images/geometry/overlapping-children-3.pdf}
    \caption{
        $m$ denotes the midpoint of $\overline{l r}$. $q'$ is the projection of $q$ onto $\overline{l r}$.
        $d$ is the distance from $q'$ to $m$.
        If $d \leq \rho$, both child clusters could contain points in the query ball.
        Otherwise, only the right child cluster could contain such points. 
    }
    \label{fig:methods:overlapping-children-3}
\end{figure}

### 2023-12-27 sections/3-datasets.tex
misc minor changes

\begin{table}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Datasets used in benchmarks.}
    \label{table:datasets:summary}
    \vskip 0.15in
    \begin{center}
        \begin{small}
            \begin{sc}
                \begin{tabular}{|l|l|l|l|}
                    \hline
                    \textbf{Dataset} & \textbf{Distance}  &\textbf{Cardinality}  & \textbf{Dimensionality}  \\
                    \hline
                    Fashion-Mnist    & Euclidean              & 60,000             & 784       \\
                    \hline
                    Glove-25         & Cosine                 & 1,183,514          & 25        \\
                    \hline
                    Sift             & Euclidean              & 1,000,000          & 128       \\
                    \hline
                    Random           & Euclidean              & 1,000,000          & 128       \\
                    \hline
                    Silva            & Levenshtein            & 2,224,640          & -         \\
                    \hline
                    RadioML          & Dynamic Time Warping   & 97,920             & 1,024     \\
                    \hline
                \end{tabular}
            \end{sc}
        \end{small}
    \end{center}
    \vskip -0.1in
\end{table}

@@

\subsection{Silva}
\label{subsec:silva}
To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the Silva dataset.
Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
We use Levenshtein distance on unaligned sequences as the distance function for this dataset.
{\color{red} TODO: Define and/or cite Levenshtein distance.}
% I think this needs another sentence but I'm not sure what else to say

\subsection{Radio ML}
\label{subsec:radioml}
To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset.
The RadioML dataset contains samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each modulation mode and SNR level~\cite{oshea2018radioml}.
Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time series of complex-valued numbers).
We used a subset of this dataset, containing only the 97,920 samples at 10dB SNR.
We use Dynamic Time Warping as the distance function for this dataset.
{\color{red} TODO: Define and/or cite Dynamic Time Warping distance.}

>>>>

\begin{table}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Datasets used in benchmarks.}
    \label{table:datasets:summary}
    \vskip 0.15in
    \begin{center}
        \begin{small}
            \begin{sc}
                \begin{tabular}{|l|l|l|l|}
                    \hline
                    \textbf{Dataset} & \textbf{Distance}  &\textbf{Cardinality}  & \textbf{Dimensionality}  \\
                    \hline
                    Fashion-Mnist    & Euclidean              & 60,000             & 784       \\
                    \hline
                    Glove-25         & Cosine                 & 1,183,514          & 25        \\
                    \hline
                    Sift             & Euclidean              & 1,000,000          & 128       \\
                    \hline
                    Random           & Euclidean              & 1,000,000          & 128       \\
                    \hline
                    Silva            & Levenshtein            & 2,224,640          & 50,000         \\
                    \hline
                    RadioML          & Dynamic Time Warping   & 97,920             & 1,024     \\
                    \hline
                \end{tabular}
            \end{sc}
        \end{small}
    \end{center}
    \vskip -0.1in
\end{table}

@@

\subsection{Silva}
\label{subsec:silva}
To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the Silva dataset.
Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
We use Levenshtein distance~\cite{levenshtein1966binary} on unaligned sequences as the distance function for this dataset.

\subsection{Radio ML}
\label{subsec:radioml}
To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset.
The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each modulation mode and SNR level~\cite{oshea2018radioml}.
Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time series of complex-valued numbers).
We used a subset of this dataset, containing only the 97,920 samples at 10dB SNR.
We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.

### 2023-12-27 sections/5-discussion.tex
misc minor changes

The questions raised by this study suggest several avenues for future work.
A comparison across more (and larder) datasets is clearly in order, as is further analysis with distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.

{\color{red} TODO: Trim this paragraph by a lot, because we never discuss compression and radii scaling in the paper and we don't want to leak too much of our thoughts on compression.}
In future work, we plan to investigate compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
When we descend some levels in the cluster tree, it takes fewer bits to encode each point in terms of the center because the radii decrease.
However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
This presents the following tradeoff:
\emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.

>>>>

The questions raised by this study suggest several avenues for future work.
A comparison across more (and larger) datasets is clearly in order, as is further analysis with distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.

% When we descend some levels in the cluster tree, it takes fewer bits to encode each point in terms of the center because the radii decrease.
% However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
% This presents the following tradeoff:
% \emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
% In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
